Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees

### 2011-06-21 paper/rmcts/paper.tex
paper in progress: figures added

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

@@

\title{Doing Better Than UCT: Rational Monte Carlo Sampling in Trees}
\author {David Tolpin \and Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

@@

\begin{abstract}

Policies for the Multiarmed Bandit (MB) proved successful for Monte Carlo
tree sampling (MCTS). However, MCTS differs from MB in that only the
final choice, rather than all arm pulls, brings a reward.

\end{abstract}

@@

\cite{Auer.ucb}

@@

\section{Doing Better Than UCB}

\section{Doing Better Than UCT}

@@

$\hat{C}_f$

@@

\section{Finite-time regret bounds}

@@

\subsection{$\varepsilon$-greedy}

Regret $R_\varepsilon$:

\begin{eqnarray}
\IE[R_\varepsilon]&=&\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{eqnarray}

@@

\begin{eqnarray}
\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\mu}{2} e^{-\mu/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \mu/4}\right)\nonumber\\
&\le&\left(\mu + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}
\end{eqnarray}

@@

\subsection{UCB}

@@

\subsection{UVB}

@@

\section{Asymptotic bounds for $\varepsilon$-greedy policies}

\section{Case study: $K-1$ and One Arms}

>>>>

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{geometry}

\geometry{margin=1.2in}

@@

\title{Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees}
\author {David Tolpin, Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

@@

\begin{abstract}
UCT, a state-of-the art algorithm for Monte Carlo tree sampling
(MCTS), is based on UCB, a sampling policy for the Multi-armed Bandit
Problem (MAB) that minimizes the accumulated regret. However, MCTS
differs from MAB in that only the final choice, rather than all arm
pulls, brings a reward, that is, the simple regret, as opposite to the
accumulated regret, must be minimized. This work introduces policies for
multi-armed bandits with lower simple regret than UCB, and an
algorithm for MCTS which combines accumulated and simple regret
minization and outperforms UCT. Finite-time and asymptotic analysis of
the policies is provided, and the algorithms are empirically compared.
\end{abstract}

@@

The simple regret of a sampling policy for the Multi-armed Bandit
Problem is the expected difference between the best expected reward
$\mu_*$ and the expected reward of the arm with the best sample mean
$\mu_j,\;\overline X_j=\max_i\overline X_i$:
\begin{equation}
\label{eq:simple-regret}
\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.

@@

\subsection{Doing better than UCB}

\begin{figure}[t]
\centering
\includegraphics[scale=1.2]{onelevel-tree.pdf}
\caption{Simple regret in MAB: the best arm}
\label{fig:onelevel-tree}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.64]{onelevel-64.pdf}
\caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. UVB}
\label{fig:onelevel-64}
\end{figure}

\subsection{Doing better than UCT}

\begin{figure}[t]
\centering
\includegraphics[scale=1.2]{twolevel-tree.pdf}
\caption{MCTS: a path to the best arm}
\label{fig:twolevel-tree}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[scale=0.64,trim=0pt 0pt 0pt 24pt,clip]{twolevel-64.pdf}
\caption{MCTS: UCT vs. UVT vs. VCT vs. Random}
\label{fig:twolevel-64}
\end{figure}

\section{Empirical Evaluation}

\subsection{Simple regret in MAB}

\begin{figure}[t]
\centering
\includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{onelevel-8-16-32-64.pdf}
\caption{Simple regret in MAB, varying number of arms}
\label{fig:onelevel-8-16-32-64}
\end{figure}

\subsection{Monte Carlo tree search}

\begin{figure}[t]
\centering
\includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{twolevel-8-16-32-64.pdf}
\caption{Monte carlo tree search, varying number of arms}
\label{fig:twolevel-8-16-32-64}
\end{figure}

\subsection{The sailing domain}

\begin{figure}[t]
  \begin{minipage}[b]{0.5\linewidth} \centering
    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
    0pt,clip]{sailing-size=8-factor=_8438.pdf}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth} \centering
    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
    0pt,clip]{sailing-size=8-nsamples=185.pdf}
  \end{minipage}
  \caption{The sailing domain, a smaller lake}
  \label{fig:sailing-8}
\end{figure}

\begin{figure}[t]
  \begin{minipage}[b]{0.5\linewidth} \centering
    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
    0pt,clip]{sailing-size=13-factor=_5625.pdf}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth} \centering
    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
    0pt,clip]{sailing-size=13-nsamples=240.pdf}
  \end{minipage}
  \caption{The sailing domain, a larger lake}
  \label{fig:sailing-13}
\end{figure}

@@

@@

\section{Derivations}

\subsection{Finite-time regret bounds}

@@

\subsubsection{$\varepsilon$-greedy}

@@

\begin{eqnarray}
\label{eq:pr-epsilon-greedy}
\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\mu}{2} e^{-\mu/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \mu/4}\right)\nonumber\\
&\le&\left(\mu + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}
\end{eqnarray}

A bound on the simple regret $R_\varepsilon$ of an
$\varepsilon$-greedy policy follows from substituting
(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):

@@

\subsubsection{UCB}

@@

\subsubsection{UVB}

@@

\subsection{Asymptotic bounds for $\varepsilon$-greedy policies}
